---
title: "IE360 HW3 SÜLEYMAN TURANCAN BURUCU"
author: "Süleyman Turancan Burucu"
date: "12/14/2020"
output: html_document
---
# A Study About the Relation Among USD/TRY Exchange Rates, Foreign Debt Amount (in million USD) and Residential Property Price Index (RPPI)

## **1. Introduction**

>In brief, the relationships between USD/TRY Exchange Rates, Foreign Debt Amount (in million USD) and Residential Property Price Index (RPPI) are examined in this study among the years 2013 - 2020 in monthly scale starting from January until December through data visualization, corelogram and Pearson's product-moment correlation.

<p>These topics are chosen exceptionally. Why I chosen these topics can be explained by Republic of Turkey's debts are generally foreign debts, i.e. borrowed in foreign currency, and this condition have been affected and have affected another economical values. To clarify, Turkish government have continued to borrow USD as a debt for as long as I've known. Also, USD against Turkish Lira exchange rate (USD/TL) is the foremost factor determining the residential prices due to the prices of iron are determined in USD globally, not in Turkish Lira (TL). Furthermore, residential properties are constructed with local materials alongwith imported materials in considerable amount. Therefore, USD/TL rate affects the foreign debt amount and RPPI.</p>

Besides, borrowing more foreign debts in USD ends up again increasing value of USD against TL. Accordingly, there is at least one vicious circle among these topics. Therefore, it can be asserted that USD/TL exchange rate affects foreign debt amount and RPPI. Yet, how much are we sure about their relationships? Can we say that the same condition holds for vice versa?</p>

If we want to reach a conclusion, we need to examine our data visually and statistically, i.e. mathematically. Therefore, I'm going to start with showing the histogram of our data.  foreign debt amount, and RPPI are shown below, respectively:</p>
>Yet, it should be kept in mind that we have been analysing only for a period, just 7 years. Therefore, we can definitely say that our calculations and results are not immutable. So, someone with different data can reach another conclusion. Also, I can reach another conclusion with another data. Therefore, conclusions below are valid only for 7 year term.

## **2. Tables and Their Explanations**
```{r, include = TRUE, echo = FALSE, warning= FALSE, error = FALSE, message = FALSE, results = "hide"}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
mydata <- data.table(date, date_year, date_month, exc, debt, ykfe)
mydata
#########################################

#Histogram of USD/TL Exchange Rate
exc_histogram <- ggplot(mydata, aes(exc, fill = date_year)) + 
  geom_histogram(bins = 30, binwidth = 0.1, alpha = 1) +
  labs(title = "Histogram of USD/TL Exchange Rate" , 
       x = "USD/TL Exchange Rate", y = "Amount", 
       fill = "Year") +
  facet_wrap(~date_year, ncol = 4) + 
  geom_density(alpha = 0.3 , fill = "black") +
  theme_minimal() + theme(legend.position = c(0.9,0.15),
                          axis.ticks.length = unit(1, "lines"),
                          legend.margin = margin(10,10,10,15,"pt"),
                          plot.title = element_text(face = "italic"),
                          rect = element_rect(fill = "grey92"),
                          panel.grid.major.y = element_line(color = "blue", size = 0.5, linetype = "dotted"),
                          legend.box.background = element_rect(fill = "grey92"))
exc_histogram

```
<p>It can be easily seen that how the bars have moved through the end of the x-axis, meaning that that from 2013 to 2020 USD/TL has increased, in general. Also, we can say that the 2018 is the year that this rate has fluctuated in large interval a lot owing to the results of degraded economical policies. Also, this condition brang forth price increases. For example, e-commerce sites have updated their imported goods' prices hourly, also some car dealer have stopped their sales until a further notice. Let's continue with another table:</p>
```{r, include = TRUE, echo = FALSE, warning= FALSE, error = FALSE, message = FALSE, results = "hide"}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
mydata <- data.table(date, date_year, date_month, exc, debt, ykfe)
mydata
#########################################

#Histogram of Foreign Debt Amount (in million USD) 
debt_histogram <- ggplot(mydata, aes(debt, fill = date_year)) +
  geom_histogram(bins = 30, alpha = 1) +
  labs(title = "Histogram of Foreign Debt Amount (in million USD)" , 
       x = "Foreign Debt Amount (in million USD)", y = "Amount",
       fill = "Year") +
  facet_wrap(~date_year, ncol = 4) + 
  geom_density(alpha = 0.3 , fill = "black") +
  theme_minimal() + theme(legend.position = c(0.9,0.15),
                          axis.ticks.length = unit(1, "lines"),
                          legend.margin = margin(10,10,10,15,"pt"),
                          plot.title = element_text(face = "italic"),
                          rect = element_rect(fill = "grey92"),
                          panel.grid.major.y = element_line(color = "blue", size = 0.5, linetype = "dotted"),
                          legend.box.background = element_rect(fill = "grey92"))
debt_histogram

```
<p>As I mentioned, Turkish government has been borrowing foreign debts for many years. Yet, bear in mind that all debts are not taken in USD, there is also another types of foreign currency Turkey have taken. Other than that, this histogram is somewhat different from the USD/TL's histogram, since it can be seen that graph above has internal consistency for each year, i.e. years are more similar to each other rather than the USD/TL plot, just because the USD graph is increasing somewhat continuously as we can see. Let's continue with another table:</p>
```{r, include = TRUE, echo = FALSE, warning= FALSE, error = FALSE, message = FALSE, results = "hide"}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
mydata <- data.table(date, date_year, date_month, exc, debt, ykfe)
mydata
#########################################
#Histogram of New Housing Price Index
ykfe_histogram <- ggplot(mydata, aes(ykfe, fill = date_year)) +
  geom_histogram(bins = 30, alpha = 1) +
  labs(title = "Histogram of Residential Property Price Index (RPPI)" , 
       x = "RPPI", y = "Amount", fill = "Year") +
  facet_wrap(~date_year, ncol = 4) + 
  geom_density(alpha = 0.3 , fill = "black") +
  theme_minimal() + theme(legend.position = c(0.9,0.15),
                          axis.ticks.length = unit(1, "lines"),
                          legend.margin = margin(10,10,10,15,"pt"),
                          plot.title = element_text(face = "italic"),
                          rect = element_rect(fill = "grey92"),
                          panel.grid.major.y = element_line(color = "blue", size = 0.5, linetype = "dotted"),
                          legend.box.background = element_rect(fill = "grey92"))
ykfe_histogram

```
<p>Lastly, we can see that the RPPI values have increased like USD/TL did, as it were the USD/TL and RPPI have traced a trend, i.e. increased more stable and alike, not like foreign debt histogram. Let's make it more clear by looking at these 3 graphs' values in one table and what I am mentioning by saying "alike":</p>
```{r, include = TRUE, echo = FALSE, warning= FALSE, error = FALSE, message = FALSE, results = "hide"}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
mydata <- data.table(date, date_year, date_month, exc, debt, ykfe)
mydata
#########################################

#Table of all exc, debt, ykfe 
allin1 <- data.table(onetwothree, date3, exc_debt_ykfe)
allin1_plot <- ggplot(allin1[, exc_debt_ykfe, by = onetwothree], 
                      aes(x = date3, y = exc_debt_ykfe)) + 
  geom_point(aes(color = allin1$onetwothree), alpha = 1) +
  labs(title = "USD/TL(scale 1/1), Foreign Debt (in million USD) (scale 1/100), 
       and RPPI (1/10)", x = "Date Starting from 01.2013 to 12.2020", y = "Amount", color = "Colors Represents: ") +
  theme_minimal() + theme(axis.text.x = element_blank(),
                          plot.title = element_text(face = "italic"),
                          legend.position = "bottom") + 
  guides(color = guide_legend(order = 1))
  
allin1_plot

```
>This graph is drawn by dividing Foreign Debt (in million USD) by 100 and RPPI by 10 and no change in the USD/TL Exchange Rate.

<p>You can easily determine that, USD/TL and RPPI values are like tracing an increasing path, which is what I trying to clarify by saying "alike" and "stable". To be more exact, we need to check the correlation between USD/TL exchange rate, foreign debt, and RPPI values. Therefore, let's look at their corelogram, a table which shows the correlations of each data one-by-one:</p>

```{r, include = TRUE, echo = FALSE, warning= FALSE, error = FALSE, message = FALSE, results = "hide"}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
mydata <- data.table(date, date_year, date_month, exc, debt, ykfe)
mydata

mydata_df <- data.frame(date, date_year, date_month, exc, debt, ykfe)
mydata_df
#########################################

#########################################
new_mydata_df <- mydata_df[,-c(1:3)]
names(new_mydata_df)[names(new_mydata_df) == "exc"] <- "USD/TL"
names(new_mydata_df)[names(new_mydata_df) == "debt"] <- "Foreign Debt Amount (in million USD)"
names(new_mydata_df)[names(new_mydata_df) == "ykfe"] <- "RPPI"
new_mydata_df
#########################################

#########################################
#Corelogram in general
ggpairs(new_mydata_df, 
        title="Corelogram Between USD/TL Exchange Rate, Foreign Debt, and RPPI",
        lower = list(continuous = wrap("smooth", alpha = 1, size=2, color = date_year, shape = 1),
                     discrete = "blank", combo="blank"),
        diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 )),
        upper = list(combo = wrap("box_no_facet", alpha=0.5)),
        ggplot2::aes(method = "pearson"))
```
<p>From that table, we can see and mathematically conclude that USD/TL and RPPI has considerably related. Predictably, we have said that USD/TL and RPPI are "stable" and "alike". At this phase, I have clarified and showed what "alikeness" is and how it can be seen by corelogram. Yet, we have said that there can be another relationship between each other,but when we look at the correlation coefficients of the each relationship, we can say that the only considerable relation is between USD/TL and RPPI, there is no relation between foreign debt vs. USD/TL, or foreign debt vs. RPPI. This condition can be explained by foreign debts are not only borrowed in USD and this situation makes the foreign debt values trace different route from USD/TL or RPPI follow.</p>

<p>If we want to check the correlations for each year, we can make another corelogram for each year whether correlation exists between USD/TL, foreign debts, and RPPI. As a reminder, we have been looking for 12 data for each year and it doesn't give the supreme and immutable relationship between them, just mathematical calculations give these results for the data sets we have:</p>

```{r, include = TRUE, echo = FALSE, warning= FALSE, error = FALSE, message = FALSE, results = "hide"}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
mydata <- data.table(date, date_year, date_month, exc, debt, ykfe)
mydata

mydata_df <- data.frame(date, date_year, date_month, exc, debt, ykfe)
mydata_df
#########################################

#########################################
new_mydata_df <- mydata_df[,-c(1:3)]
names(new_mydata_df)[names(new_mydata_df) == "exc"] <- "USD/TL"
names(new_mydata_df)[names(new_mydata_df) == "debt"] <- "Foreign Debt Amount (in million USD)"
names(new_mydata_df)[names(new_mydata_df) == "ykfe"] <- "RPPI"
new_mydata_df
#########################################

#########################################
#Correlogram of USD/TL Exchange Rate, Foreign Debt, and RPPI for Each Year
ggpairs(new_mydata_df,
        title="Correlogram  Between USD/TL Exchange Rate, Foreign Debt, and RPPI for Each Year",
        lower = list(continuous = wrap("smooth", alpha = 0.8, size=2, color = date_year, shape = 1),
                     discrete = "blank", combo="blank"),
        diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 )),
        upper = list(combo = wrap("box_no_facet", alpha=0.5)),
        ggplot2::aes(colour=date_year,method = "pearson")) 
```
<p>Here we can observe that the USD/TL and RPPI are positively correlated especially for 2013 and 2015 and positively correlated in general. As I said, we can examine our data in various aspects and find different results. Also, even we find unimportant correlation coefficients among foreign debt - RPPI and foreign debt - USD/TL, in some of the years our program has caught some correlation but not so important. Presumably, this condition results from having only 12 data for each year, and the correlation might not be found if we examined the each year by lookin their daily values. Yet, some of the values are not calculated in a daily basis, unfortunately.</p>
<p>If we want to see the results of Pearson's product-moment correlation, which is same with what we have found, showing the correlations between USD/TL exchange rate, foreign debt and RPPI more formally.</p>

<p>Here, the results for the correlation between USD/TL exchange rate and foreign debt:</p>

```{r, include = TRUE, echo = FALSE, warning= FALSE}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
print(cor.test(exc, debt))
```
<p>Here, the results for the correlation between USD/TL exchange rate and RPPI is shown:</p>

```{r, include = TRUE, echo = FALSE, warning= FALSE}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
print(cor.test(exc, ykfe))
```
<p>Here, the results for the correlation between foreign debt and RPPI is shown:</p>
```{r, include = TRUE, echo = FALSE, warning= FALSE}
setwd("D:/Lectures/(2020-2021) Lectures Junior/IE360/Homeworks/2")

library(ggplot2)
library(knitr)
library(markdown)
library(data.table)
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)
library(GGally)
library(lubridate)

#######################
date <- c("2013-01","2013-02","2013-03","2013-04","2013-05","2013-06","2013-07","2013-08","2013-09","2013-10","2013-11","2013-12",
          "2014-01","2014-02","2014-03","2014-04","2014-05","2014-06","2014-07","2014-08","2014-09","2014-10","2014-11","2014-12",
          "2015-01","2015-02","2015-03","2015-04","2015-05","2015-06","2015-07","2015-08","2015-09","2015-10","2015-11","2015-12",
          "2016-01","2016-02","2016-03","2016-04","2016-05","2016-06","2016-07","2016-08","2016-09","2016-10","2016-11","2016-12",
          "2017-01","2017-02","2017-03","2017-04","2017-05","2017-06","2017-07","2017-08","2017-09","2017-10","2017-11","2017-12",
          "2018-01","2018-02","2018-03","2018-04","2018-05","2018-06","2018-07","2018-08","2018-09","2018-10","2018-11","2018-12",
          "2019-01","2019-02","2019-03","2019-04","2019-05","2019-06","2019-07","2019-08","2019-09","2019-10","2019-11","2019-12")

date_year <-rep(c("2013","2014","2015","2016","2017","2018","2019"), each = 12)

date_month <- rep(c("1","2", "3", "4", "5", "6","7","8","9","10","11","12"), times = 7)

exc <- c(1.764,1.770,1.807,1.796,1.823,1.895,1.930,1.955,2.017,1.990,2.022,2.058,
         2.217,2.213,2.218,2.127,2.091,2.116,2.119,2.158,2.204,2.258,2.234,2.288,
         2.328,2.455,2.584,2.648,2.646,2.701,2.695,2.846,3.003,2.930,2.871,2.917,
         3.007,2.941,2.892,2.835,2.927,2.917,2.958,2.963,2.960,3.068,3.267,3.489,
         3.735,3.672,3.666,3.654,3.564,3.519,3.560,3.512,3.468,3.662,3.879,3.848,
         3.772,3.778,3.881,4.054,4.414,4.628,4.748,5.730,6.367,5.859,5.373,5.306,
         5.369,5.262,5.442,5.736,6.049,5.812,5.672,5.625,5.713,5.785,5.731,5.838)

debt <- c(2791.85,1011.50,837.71,830.14,1086.38,514.98,719.41,318.86,657.94,884.97,525.06,565.64,
          2585.33,2150.60,1022.88,795.69,488.00,595.09,690.71,494.48,801.47,1113.18,413.16,785.99,
          914.56,491.91,3566.03,929.23,475.46,356.43,761.28,433.78,902.88,556.65,329.35,589.21,
          1286.80,490.60,2146.26,674.78,568.05,414.52,575.99,448.89,2712.13,585.49,498.12,580.89,
          701.17,2092.18,691.36,744.05,679.33,418.79,2812.93,536.39,1283.88,795.23,694.16,1365.45,
          1816.53,496.74,2353.13,3007.68,740.14,378.75,539.10,586.99,817.22,1851.03,523.01,623.63,
          524.08,577.14,2255.75,2115.56,605.55,685.15,564.24,923.08,742.66,1104.70,2078.13,909.84)

ykfe <- c(56.400,57.500,58.200,58.700,59.300,60.000,60.800,61.400,61.900,62.200,63.100,63.600,
          63.600,64.300,65.100,65.900,66.400,67.300,68.200,69.100,69.800,70.400,71.300,71.800,
          72.900,74.400,75.500,76.500,77.300,78.500,79.500,80.300,80.700,81.400,82.200,83.300,
          84.700,86.000,86.600,87.500,88.600,89.600,90.500,91.200,91.500,92.400,93.100,93.700,
          95.000,96.600,97.800,98.700,99.200,99.800,100.200,101.200,101.700,102.700,103.200,103.900,
          104.000,105.200,106.200,107.700,108.600,109.200,109.200,108.700,108.300,108.000,108.800,110.300,
          109.500,109.400,109.300,110.200,110.100,112.000,114.200,115.600,116.900,118.400,120.100,122.100)

onetwothree <- rep(c("USD/TL Exchange Rate","Foreign Debt Amount (in million USD)","RPPI"), each = 84)

date3 <- rep(c(date), times = 3)

exc_debt_ykfe <- c(exc, debt/100, ykfe/10)


#######################

#########################################
print(cor.test(debt, ykfe))
```

## **3. Conclusion**
<p>As a consequence, the relationships between USD/TRY Exchange Rates, Foreign Debt Amount (in million USD) and Residential Property Price Index (RPPI) are examined in this study among the years 2013 - 2020 in monthly scale starting from January until December through data visualization, corelogram and Pearson's product-moment correlation. Then, it was found that USD/TRY Exchange Rates and Residential Property Price Index (RPPI) are found to be correlated with each other, seen from both the corelogram and the Pearson's product-moment correlation calculations both with same results, deduced from having large t-value and small p-value only at the USD/TL exchange rate and RPPI's correlation calculation. Some of the specific years' data are found to be correlated in itself;however, this situation supposably resulted from the data amount, i.e. if we have more data then we might confidently asserted that these data were related with each other.</p>

The .Rmd file can be found by clicking here[]

## **4. Resources**
* [EVDS](https://evds2.tcmb.gov.tr/index.php?/evds/serieMarket/collapse_2/5868/DataGroup/turkish/bie_rktufey/)</p>
* [Stackoverflow](https://stackoverflow.com/)</p>
* [A Brief Introduction to ggpairs()](https://www.blopig.com/blog/2019/06/a-brief-introduction-to-ggpairs/)</p>
* [tidyverse.org](https://ggplot2.tidyverse.org/)</p>
* [statmethods.net](https://www.statmethods.net/)</p>
* [RStudio Community](https://community.rstudio.com/)</p>
* [r-graph-gallery.com](https://www.r-graph-gallery.com/)</p>
* [Datacamp](https://www.datacamp.com/)
